


\documentclass[proquest]{uwthesis}[2020/02/24]
\usepackage{graphicx}
\usepackage{caption}
\usepackage{listings}
\usepackage{listings}
\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  frame=single,
  breaklines=true,
  %postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
}
\usepackage{subcaption}
\usepackage{blindtext}
\usepackage{hyperref}
\usepackage[section]{placeins} % force figures into place
\usepackage{setspace} % to decrease line spacing locally
\usepackage{amsmath} % for better math
\usepackage[shortcuts]{extdash} % for dash without line break
% Setup fnsymbols with both numbers and symbols for author list footnotes:
\usepackage[symbol*]{footmisc}
\DefineFNsymbolsTM{author_list}{
  1 1
  2 2
  3 3
  \textdagger    \dagger
  \textasteriskcentered *
  \textdaggerdbl \ddagger
  \textsection   \mathsection
  \textbardbl    \|%
  \textparagraph \mathparagraph
}
\setfnsymbol{author_list}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdfpagemode=FullScreen,
}

% With so many floats, vertical alignment becomes stupid:
\raggedbottom


\setcounter{tocdepth}{1}  % Print the chapter and sections to the toc
 

% ==========   Local defs and mods
%

% --- sample stuff only -----
% These format the sample code in this document

\usepackage{alltt}  % 
\newenvironment{demo}
  {\begin{alltt}\leftskip3em
     \def\\{\ttfamily\char`\\}%
     \def\{{\ttfamily\char`\{}%
     \def\}{\ttfamily\char`\}}}
  {\end{alltt}}
 
% metafont font.  If logo not available, use the second form
%
% \font\mffont=logosl10 scaled\magstep1
\let\mffont=\sf
% --- end-of-sample-stuff ---





%%%% Shortcuts %%%%%

\newcommand{\natlog}{\text{ln}}
\newcommand{\Asn}{\text{Asn}}
\newcommand{\UAsn}{\text{U-\textsuperscript{13}C Asn}}
\newcommand{\Flin}{J_{\text{in}}}
\newcommand{\Flout}{J_{\text{out}}}
\newcommand{\Flsyn}{J_{\text{syn}}}
\newcommand{\Flprot}{J_{\text{prot}}}
\newcommand{\hCi}{\textsuperscript{13}C}
\newcommand{\hNi}{\textsuperscript{15}N}
\newcommand{\NAD}{NAD\textsuperscript{+}}
\newcommand{\NADP}{NADP\textsuperscript{+}}





 


\begin{document}
 
% ==========   Preliminary pages
%
% ( revised 2012 for electronic submission )
%

\prelimpages
 
%
% ----- copyright and title pages
%
\Title{Tools to investigate aspartate limited proliferation\\ of cancer cells}
\Author{Kristian Davidsen}
\Year{2024}
\Program{Molecular and cellular biology}

\Chair{Lucas B Sullivan}{}{Fred Hutch Cancer Center, Human Biology}
\Signature{Arvind R Subramaniam}
\Signature{Dana L Miller}

\copyrightpage
\titlepage  


%
% ----- abstract
%





\setcounter{page}{-1}
\abstract{%
Cellular aspartate levels are tightly linked to cell proliferation rate.
But how aspartate limitation decreases proliferation is unknown and could be related to its role as an amino acid as well as a substrate in multiple metabolic reactions.
In this dissertation, the role of each of the metabolic fates of aspartate is interrogated.
Surprisingly, it was not possible to rescue proliferation during aspartate limitation by providing the fates of aspartate, neither were there any evidence of insufficiency for protein synthesis.
It is hypothesized that nucleotide imbalance is responsible and that this would elude efforts to rescue proliferation through complementation.
Nevertheless, these results prompted the development of better tools to measure aspartate and the consequences of aspartate limitation.
Towards that end, an aspartate biosensor was developed and tested, enabling increased throughput of non-invasive aspartate measurements under relevant growth conditions.
Also, a highly accurate method for measuring tRNA aminoacylation using sequencing was developed.
This tRNA-Seq method was tested on human tRNAs with a titration of charge levels, its measurements were determined to be quantitative and with enough accuracy and precision to detect small ($\sim$5\%) effects on tRNA charge.
}
 
%
% ----- contents & etc.
%
\tableofcontents
\listoffigures
\listoftables
 
% Generally, better to write in a way such that no glossary is necessary
%\include{sections/glossary.tex}

% end of the preliminary pages
 
 
 
%
% ==========      Text pages
%

\textpages

\include{sections/ackn.tex}

%»Det værste ved ikke at lave noget er, at man aldrig ved, hvornår man er færdig.«\\
%Storm P



% General introduction to aspartate metabolism,
% starting with the bascis then moving into aspartate limitation
% and the relevant papers:
\include{sections/chap1.tex} % Aspartate in cellular metabolism

% All the negative data collected throughout the years:
\include{sections/chap2.tex} % Molecular mechanisms of aspartate limitation

% Asp-sensor manuscript:
\include{sections/chap3.tex} % An engineered biosensor enables dynamic aspartate measurements in living cells

% Describe all things relevant to understand tRNA-Seq method:
%KD: Hell! Too lazy to write this crap.
%\include{sections/chap4.tex} % Introduction to tRNA biology

% tRNA-Seq manuscript:
\include{sections/chap5.tex} % Robust method for measuring aminoacylation through tRNA-seq

% Future directions:
\include{sections/chap6.tex}






%
% ==========   Bibliography
%
% \nocite{*}   % include everything in the uwthesis.bib file
\bibliographystyle{plain}
\bibliography{uwthesis}
%
% ==========   Appendices
%
\appendix
\raggedbottom\sloppy
 
% ========== Appendix A


\include{appendices/app_ch1}
\include{appendices/app_ch2}
\include{appendices/app_ch3}
\include{appendices/app_ch5}

% \include{appendices/sapp}







 




\end{document}
